A Test of the Frequency Independence Assumption of Power System Coefficients used in Geomagnetically Induced Current Estimates

### 2019-08-01 paper-submitted-2.tex
Updates from Overleaf

\begin{keypoints}
\item Power system coefficients derived from a time series of GIC and geoelectric field measurements are frequency dependent.
-\item Data-derived power system coefficients may differ significantly with those computed using power system configuration information.
\end{keypoints}

@@

The PE and MSE ratios are shown in Table~\ref{resultstable}. Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~\ref{resultstable}, i.e., Model~1 has the lowest PE and Model~4 has the highest. Note that the large errors visible at the start and end of the time intervals are due to the transients discussed in Section~\ref{section:Model_Evaluation} and the first and last 10~minutes of each interval were omitted from metrics calculations.

@@

Figure~\ref{Z} shows that the frequency domain transfer functions for Model~2 vary by an order of magnitude on timescales of 60~s to 12~h (that for Model~1 is constant by definition). For this system, GIC is more sensitive to $E_y$ than $E_x$ due to $b(\omega) > a(\omega)$, but note that $E_y(t)$ has less variability than $E_x(t)$ - the variance in $E_x(t)$ is $\sim$5x larger than $E_y(t)$ over the entire dataset (this is also visible in Figure~\ref{sample}).

@@

Both Model~1 and~2 predict a maximum GIC when the electric field angle is significantly different from the angle of the 187~kV line. As discussed in the following section, this can be caused by other parts of the power network (the details of which are unavailable) making significant contributions to GIC and variations in ground conductivity. 
Figure~\ref{Z_MT} and~\ref{Phi_MT} show the parameters for $\mathbf{Z}$ in Model~3. Note that transfer function shown in Figure~\ref{Z_MT} differs significantly from that of a layered slab conductivity model for which $|Z_{xy}|=|Z_{yx}|$. That electric and magnetic field data from Memanbetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$. When using the \cite{Fujii2015} transfer function for Model~3, the PE was significantly lower (0.58 vs. 0.78). Although not critical to the conclusions in this work, a preliminary investigation suggests several explanations. We have found that when using continuous one-month intervals prior to 2006 to compute $\mathbf{Z}$, our calculated transfer function matches the \cite{Fujii2015} transfer function to within the error bars except below periods of 100~s, which is likely explained by the issues with the electric field response in the 1-second-cadence measurements discussed in \cite{Fujii2015} and the fact that we did not use a remote reference in our calculations. After 2006 (January through May tested) our calculated transfer function magnitudes differ significantly from that in \cite{Fujii2015} - they are shifted downward by approximately 10-50\%. In addition, we have found that the estimated transfer function is sensitive to whether or not segment averaging was used. Due to the large gaps between and the varying length of continuous segments of data used in this work, transfer function parameters were computed using one-day segments of data and then averaged. We have found that if a continuous set of measurements over approximately 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% larger than if 21 transfer functions were computed and then averaged. This result is unexpected, and results from extensive verification tests of our MT calculations along with the fact that this effect has observed by both authors using two different MT calculation codes on different data sets (the other code used was LEMI MT V1.4), suggests the explanation is not due to an error in the code used to compute the transfer functions in this work. We expect to consider this in a future investigation as it suggests that stack averaging can be a significant source of transfer function bias and expect that the explanation involves non-stationarity in the noise.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_GIC_predictions-MeanModel-2006-08-06-v0-o0.pdf}
  \caption{Out-of-sample model predictions (top) and predictions errors for each model in a selected 1-day interval. Note that the scales for the error time series differ from that for the top panel so that the error amplitudes are scaled up by a factor of 2.}
  \label{predictions}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_model_summary_aobo_histograms-options-1-v0-o0.pdf}
  \caption{Distribution of coefficients in Model~1. Each set of coefficients was computed 35 times using one day of 1-second cadence data. The circle or dot on the horizontal axis indicates the average of a histogram: $a_o$ = 73 A/(V/km) and $b_o$ = -72 A/(V/km).}
  \label{histogram}
\end{figure}

>>>>

\begin{keypoints}
\item Power system coefficients derived from a time series of GIC and geoelectric field measurements are frequency dependent.
\item A GIC model with frequency-dependent system coefficients provides significantly better estimates of GIC than a frequency independent model.
\item Data-derived power system coefficients may differ significantly with those computed using power system configuration information.
\end{keypoints}

@@

The PE and MSE ratios are shown in Table~\ref{resultstable}. Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~\ref{resultstable}, i.e., Model~1 has the lowest PE and Model~4 has the highest. The large errors visible at the start and end of the time intervals are due to the transients discussed in Section~\ref{section:Model_Evaluation} and the first and last 10~minutes of each interval were omitted from metrics calculations.

@@

Figure~\ref{Z} shows that the frequency domain transfer functions for Model~2 vary by an order of magnitude on timescales of 60~s to 12~h (that for Model~1 is constant by definition). For this system, GIC is more sensitive to $E_y$ than $E_x$ due to $b(\omega) > a(\omega)$, but $E_y(t)$ has less variability than $E_x(t)$ - the variance in $E_x(t)$ is $\sim$5x larger than $E_y(t)$ over the entire dataset (this is also visible in Figure~\ref{sample}).

@@

Both Model~1 and~2 predict a maximum GIC when the electric field angle is significantly different from the angle of the 187~kV line. As discussed in the following section, this can be caused by other parts of the power network (the details of which are unavailable) making significant contributions to GIC and variations in ground conductivity. 
Figure~\ref{Z_MT} and~\ref{Phi_MT} show the parameters for $\mathbf{Z}$ in Model~3. Note that transfer function shown in Figure~\ref{Z_MT} differs significantly from that of a layered slab conductivity model for which $Z_{xy}=-Z_{yx}$ and $Z_{xx}=Z_{yy}=0$. That electric and magnetic field data from Memanbetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$. When using the \cite{Fujii2015} transfer function for Model~3, the PE was significantly lower (0.58 vs. 0.78). Although not critical to the conclusions in this work, a preliminary investigation suggests several explanations. We have found that when using continuous one-month intervals prior to 2006 to compute $\mathbf{Z}$, our calculated transfer function matches the \cite{Fujii2015} transfer function to within the error bars except below periods of 100~s, which is likely explained by the issues with the electric field response in the 1-second-cadence measurements discussed in \cite{Fujii2015} and the fact that we did not use a remote reference in our calculations. After 2006 (January through May tested) our calculated transfer function magnitudes differ significantly from that in \cite{Fujii2015} - they are shifted downward by approximately 10-50\%. In addition, we have found that the estimated transfer function is sensitive to whether or not segment averaging was used. Due to the large gaps between and the varying length of continuous segments of data used in this work, transfer function parameters were computed using one-day segments of data and then averaged. We have found that if a continuous span of measurements over 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% larger than if 21 transfer functions were computed and then averaged. This result is unexpected, and results from extensive verification tests of our MT calculations along with the fact that this effect has observed by both authors using two different MT calculation codes on different data sets (the other code used was LEMI MT V1.4), suggests the explanation is not due to an error in the code used to compute the transfer functions in this work. This result suggests that stack averaging can be a significant source of transfer function bias and we expect that the explanation involves variations in the signal-to-noise between segments.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_GIC_predictions-MeanModel-2006-08-06-v0-o0.pdf}
  \caption{Out-of-sample model predictions (top) and predictions errors for each model in a selected 1-day interval. The scales for the error time series differ from that for the top panel so that the error amplitudes are scaled up by a factor of 2.}
  \label{predictions}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_model_summary_aobo_histograms-options-1-v0-o0.pdf}
  \caption{Distribution of coefficients in Model~1. Each set of coefficients was computed 35 times using one day of 1-second cadence data. The circle or dot on the horizontal axis indicates the average of a histogram and $\overline{a}_o$ = 73 A/(V/km) and $\overline{b}_o$ = -72 A/(V/km).}
  \label{histogram}
\end{figure}
